\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Local Adaptive Optimization of Time Step}

### 2013-10-29 report/article.tex
Added references for various simple concepts

The total force of a spring is the sum of the spring force $F_{h,d}$
calculated by using Hooke's Law and the damping force $F_{C,d}$
using a simple damping method \todo{Add reference (wiki):
    \url{http://en.wikipedia.org/wiki/Harmonic_oscillator}}: \begin{align}
\nonumber l_d &= p_{l,d} - p_{r,d} \\ \label{eq:hooke} F_{h,d} &= -k_d x_d
= -k_d (l_d - l_{0,d}) \\ \label{eq:damping} F_{C,d} &= -C_d \frac{\partial
x_d}{ \partial t} = - C_d \frac{\left(v_{l,d} - v_{r,d}\right) l_d}{\abs{l_d}}
\\ F_d &= F_{h,d} + F_{C,d} \end{align} \todo{Maybe add direction correction:
    $F_d \cdot (-l_d / \abs{l_d})$} in which $p_{l,d}$ and $p_{r,d}$ are the
positions of the left and right endpoints of the spring $d$ respectively, and
$v_{l,d}$ and $v_{r,d}$ are the velocities of the left and right endpoint
vertices of the spring $d$ respectively, and $F_d$ is the total force of
spring $d$.

@@

We now know how to calculate the force of the spring system and apply it
to the vertices. In order to calculate the derrivatives of the system
(step 3) we use Newton's second law of motion \todo{Add reference,
    \url{http://en.wikipedia.org/wiki/Newton's_laws_of_motion}} to get the
difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex $v$
using a given time step size $\Delta t$: \begin{align} \label{eq:delta_u}
\Delta u_v &= \frac{\partial u_v}{\partial t} \Delta t = \frac{F_v}{m_v}
\Delta t \\ \label{eq:delta_p} \Delta p_v &= u \Delta t \end{align}

As mentioned in the beginning of this section we have chosen to use the
explicit Euler method defined below in equation \ref{eq:explicit_euler}
to calculate the time ``integration'' \todo{Add reference,
    \url{http://en.wikipedia.org/wiki/Euler_method}} when stepping through
the simulation. Given a variable $x$ as a function of the time $t$, the
variable $x$ at the time $t + \Delta t$ is defined as: \begin{align}
\label{eq:explicit_euler} x(t+\Delta t) &= x(t) + \Delta t \frac{\partial
x(t)}{\partial t} \end{align} Applying this method for the timestepping
of our simulation we get the two following position and velocity update
equations using equations \ref{eq:delta_u} and \ref{eq:delta_p}: \begin{align}
p_v(t+\Delta t) &= p_v(t) + \Delta p_v \\ u_v(t+\Delta t) &= u_v(t) + \Delta
u_v \end{align} We use these equations to perform step 4 of the simulation.

>>>>

The total force of a spring is the sum of the spring force $F_{h,d}$
calculated by using Hooke's Law (see \cite[p.~439]{Young:2010})
and the damping force $F_{C,d}$ using a simple damping method (see
\cite[p.~457]{Young:2010}): \begin{align} \nonumber l_d &= p_{l,d} -
p_{r,d} \\ \label{eq:hooke} F_{h,d} &= -k_d x_d = -k_d (l_d - l_{0,d}) \\
\label{eq:damping} F_{C,d} &= -C_d \frac{\partial x_d}{ \partial t} = - C_d
\frac{\left(v_{l,d} - v_{r,d}\right) l_d}{\abs{l_d}} \\ F_d &= F_{h,d} +
F_{C,d} \end{align} \todo{Maybe add direction correction: $F_d \cdot (-l_d
/ \abs{l_d})$} in which $p_{l,d}$ and $p_{r,d}$ are the positions of the
left and right endpoints of the spring $d$ respectively, and $v_{l,d}$ and
$v_{r,d}$ are the velocities of the left and right endpoint vertices of the
spring $d$ respectively, and $F_d$ is the total force of spring $d$.

@@

We now know how to calculate the force of the spring system and apply it to
the vertices. In order to calculate the derrivatives of the system (step 3) we
use Newton's second law of motion (see \cite[p.~112-115]{Young:2010}) to get
the difference in speed $\Delta u_v$ and position $\Delta p_v$ of each vertex
$v$ using a given time step size $\Delta t$: \begin{align} \label{eq:delta_u}
\Delta u_v &= \frac{\partial u_v}{\partial t} \Delta t = \frac{F_v}{m_v}
\Delta t \\ \label{eq:delta_p} \Delta p_v &= u \Delta t \end{align}

As mentioned in the beginning of this section we have chosen to use the
explicit Euler method defined below in equation \ref{eq:explicit_euler}
to calculate the time ``integration'' \cite{Flaherty} when stepping through
the simulation. Given a variable $x$ as a function of the time $t$, the
variable $x$ at the time $t + \Delta t$ is defined as: \begin{align}
\label{eq:explicit_euler} x(t+\Delta t) &= x(t) + \Delta t \frac{\partial
x(t)}{\partial t} \end{align} Applying this method for the timestepping
of our simulation we get the two following position and velocity update
equations using equations \ref{eq:delta_u} and \ref{eq:delta_p}: \begin{align}
p_v(t+\Delta t) &= p_v(t) + \Delta p_v \\ u_v(t+\Delta t) &= u_v(t) + \Delta
u_v \end{align} We use these equations to perform step 4 of the simulation.
